---
title: "Setting Up"
author: "CY"
date: "09/10/19"
output:
  html_document:
    theme: spacelab
    toc: yes
    toc_float: yes
---

## Resources
- [R for data science](https://r4ds.had.co.nz/)
- [R cheatsheets](https://www.rstudio.com/resources/cheatsheets/)
- [RStudio keyboard shortcuts](https://support.rstudio.com/hc/en-us/articles/200711853-Keyboard-Shortcuts)


## Introduction
We will get started using R, Rstudio, and GitHub. All of our work will be done in RStudio and uploaded to the group GitHub repository.  


## GitHub and Git
[**GitHub**](www.github.com) is a file storage and management site used mostly by programmers. Programmers upload code to repositories (folders) and make it publicly available.  

**Git** is software that is used for version control. It tracks changes and is especially helpful when multiple people are collaborating on a project; if a mistake is made it can be easily tracked back and retrieved. Both can be integrated with RStudio.  

Most programmers use both **GitHub** and **Git** to manage their projects. However, **Git** is glitchy and hard to work with- especially in a classroom setting. If you want to try and use **Git**, then follow the directions [here](https://jmledford3115.github.io/datascibiol/setup.html).  

## GitHub Desktop
The easiest way to manage your GitHub repositories is to use [GitHub desktop](https://desktop.github.com/). This is a  helpful tool that will allow you to manage all of your files. Highly recommended!  

## Create a GitHub account
Please create a GitHub account and send me your username :) 

I have made a repository for our group titled: [ATS-Masters/R tutorial](https://github.com/ats-master/r_tutorial/tree/Notes). Once you have accepted the invitation to join the repository, you should be able to login to GitHub and see this organization or click on the link above. You should see a directory associated with your login name. Feel free to upload your working examples here.


## RStudio
RStudio is an interactive working environment for R. When you first open it, you should see your screen divided into four quadrants. Basically they display different information and are helpful as we work on code.  

You need to make a folder on your computer where you will keep all of your work. You will need frequent access to the folder, so I recommend putting it in an easy to find space. I put mine on the desktop.  

Please run the following code to check your current working directory.
```{r}
getwd()
```

If you find that you are not in the folder that you created, please navigate to it: Session>Set Working Directory>Choose Directory.

Re-run the code below to confirm.
```{r}
getwd()
```

## RMarkdown
RMarkdown is one of many types of documents that can be created in RStudio. It is an extremely helpful way to build code because it will display the code, its output, and any text that you wish to add.  

In RStudio, open a new markdown document: **File>New File>R Markdown**  

There is some generic information at the start of the file that we don't need. Delete lines 12 through 31 to clean things up.  

In this file insert a new R code chunk by clicking on the green `insert` tab and selecting R.  
Our mac shortcut is option + command + i  


### Install the tidyverse
One strength of R is that there are thousands of add-on packages that perform specialized functions. The packages are referred to as libraries.  Libraries need to be installed in order to work, and whenever you update R you also need to update all of the packages that you use. We will be using the library "tidyverse" regularly.
```{r eval=FALSE, include=TRUE}
install.packages("tidyverse")
```

## Push files to GitHub
You can use your repository to store the files. All you need to do is go back to your GitHub directory in your web browser and click `Upload files`. Once there, you can drag and drop any files directly into your repository. But, in order for the upload to complete you need to add a short message in the box beneath `Commit changes`. Once the message is added, click `Commit changes` and your upload will be complete.


